\mytitle

### 2021-12-16 paper.tex
reformulation

Several modern CPU architectures support instruction\-/level
parallelism (ILP), which allows for the execution of multiple
instructions at the same time.  To make use of ILP, one could use
Estrin's scheme \cite{estrin} to rewrite the function as
%
\begin{align}
\begin{split}
  \ln(1+x) \approx{}& x + \left(\frac{x}{3}-\frac{1}{2}\right)
    y \\ &+\left[\left(\frac{x}{7}-\frac{1}{6}\right)
    y+\frac{x}{5}-\frac{1}{4}\right] z,
\end{split}\label{eq:estrin}%
\end{align}
%
where $y=x^2$ and $z=y^2$.  This form requires 8 floating\-/point
multiplications and 6 additions, i.e.\ more multiplications than
Horner's scheme.  However, in Estrin's form several terms can be
executed in parallel, leading to a potential speed-up if the cost of
the additional multiplications is smaller than the gain by the
parallel execution.  A C implementation using Estrin's scheme could
read:
%
\lstinputlisting[language=C]{src/log1p_estrin.c}

>>>>

There are several different possibilities to reduce the runtime of
Horner's method by rewriting the series expansion of $\ln(1+x)$ such
that the more terms can be calculated in parallel, while the number
floating\-/point operations is still kept small.  One possibility is
to split the r.h.s.\ of Eq.~\eqref{eq:horner} into two distinct sums,
one containing the even powers and the other containing the odd powers
of $x$.  Both sums can then be calculated in parallel.  Another
possibility is to rewrite r.h.s.\ of Eq.~\eqref{eq:taylor} using
Estrin's scheme \cite{estrin},
%
\begin{align}
\begin{split}
  \ln(1+x) \approx{}& x + \left(\frac{x}{3}-\frac{1}{2}\right)
    y \\ &+\left[\left(\frac{x}{7}-\frac{1}{6}\right)
    y+\frac{x}{5}-\frac{1}{4}\right] z,
\end{split}\label{eq:estrin}%
\end{align}
%
where $y=x^2$ and $z=y^2$.  This form requires 8 floating\-/point
multiplications and 6 additions, i.e.\ one more multiplication than
Horner's scheme.  However, in Estrin's form several terms can be
executed in parallel, leading to a potential speed-up if the cost of
the additional multiplications is smaller than the gain by the
parallel execution.  A C implementation using Estrin's scheme could
read:
%
\lstinputlisting[language=C]{src/log1p_estrin.c}
